EDIpack 2.0: A generic and interoperable Lanczos-based solver for quantum impurity problems

### 2025-05-20 01_intro.tex
Read introduction ad install (LC)

A key development feature in \NAME, unlocked by the introduction of Python and C/C++ API, 
is the direct  interfaces to the TRIQS (Toolbox for Research on Interacting Quantum Systems) and
W2Dynamics software suites, which are two of the most important and used computational tools in
the field. \NAME seamlessly slots in as an alternative solver to the native Quantum Monte Carlo 
implementations, greatly extending the parameter (and especially temperature) range reachable by 
the simulation while minimizing the cfor the end user.
At the same time, this integration provides \NAME users with access to the
extensive data handling and analysis tools in the TRIQS ecosystem. 
Overall, interoperability extends the cumulative reach and trustworthiness of all these platforms,
enhancing reproducibility and cross-validation of numerical results, with the ultimate aim of achieving standardization for quantum impurity solvers.

>>>>

A key development feature in \NAME, unlocked by the introduction of Python and C/C++ API, 
is the direct interfaces to the TRIQS (Toolbox for Research on Interacting Quantum Systems) and
W2Dynamics software suites, which are two of the most important and used computational tools in
the field. \NAME seamlessly slots in as an alternative solver to the native Quantum Monte Carlo 
implementations, greatly extending the parameter (and especially temperature) range reachable by 
the simulation while minimizing the coding requirements for the end user.
At the same time, this integration provides \NAME users with access to the
extensive data handling and analysis tools in the TRIQS ecosystem. 
Overall, interoperability extends the cumulative reach and trustworthiness of all these platforms,
enhancing reproducibility and cross-validation of numerical results, with the ultimate aim of a
chieving standardization for quantum impurity solvers along the guiding principles of initiatives such as FAIRmat.

### 2025-05-20 02_install.tex
Added some Julia comments
Read introduction ad install (LC)

\section{Installation}\label{SecInstall}
The installation of \NAME is available through CMake which ensures
multi-platforms compabitility and dependencies resolution.  
The software builds into two distinct libraries.
The main one is {\tt  libedipack.a} which, alongside the generated Fortan
modules, wraps the \NAME software possibly including support for  
inequivalent impurities.
A second dynamic library, {\tt libedipack\_cbinding.so}, enables interoperability through
specific bindings to the C/C++ programming language.

\subsection{Structure}\label{sSecInstallStructure}
\NAME is a modular library organized into three main components. At its 
core is the exact diagonalization solver, which provides the primary 
computational engine. Building on this, the library includes 
{\tt EDIpack2ineq}, an extension for handling multiple inequivalent 
impurity problems. Finally, a Fortran-C/C++ interface is provided for 
seamless integration with external software and the development of 
additional APIs.

@@

\subsection{Dependencies}\label{sSecInstallDependencies}
\NAME essentially depends on two external libraries.
\begin{itemize}
\item {\bf SciFortran}: an open-source Fortran library to support
  mathematical and scientific software development. 
\item {\bf MPI} (optional): a distributed memory parallel communication layer with support to modern Fortran compiler.
\end{itemize}

@@

\subsection{Build and Install}\label{sSecInstallBuildInstall}
\subsubsection{Source}
The software can be installed from source as follows. The source can
be retrieved directly from its GitHub repository, for instance using:
\begin{lstlisting}[style=mybash]
git clone https://github.com/edipack/EDIpack2.0 EDIpack2
\end{lstlisting}
Then, assuming to be in the software directoru, a conventional
out-of-source building is performed using two different compilations
backends.

@@

\noindent
The CMake configurations can be further tuned using the following variables:
\begin{center}
\begin{tabular}{ l|l|l } 
 \hline
  {\bf Option}               & {\bf Scope} & {\bf Value} \\
  \hline
  -D{\bf PREFIX}          & Install directory  & $\sim$/opt/EDIpack2/TAG/PLAT/BRANCH\\
  -D{\bf USE\_MPI}       & MPI support  &  True/{\color{red}False}\\
  -D{\bf WITH\_INEQ}   & Inequivalent impurities support & {\color{red}True}/{False}\\
  -D{\bf VERBOSE}      & Verbose CMake output & {\color{red}True}/{False}\\ 
  -D{\bf BUILD\_TYPE} & Compilation flags & {\color{red}RELEASE}/TESTING/DEBUG/AGGRESSIVE \\
 \hline
\end{tabular}
\end{center}

The default target builds and install either the main library and the C-binding.
However, a specific building for each library is available specifying
the required target. A recap message is printed at the end of the
CMake configuration step.

\subsubsection{Anaconda}
As an alternative we provide for both Linux and OSx systems
installation through Anaconda packages into a virtual
environment containing Python ($>3.10$).

The Conda package installation procedure reads:
\begin{lstlisting}[style=mybash]
conda create -n edipack
conda activate edipack
conda install -c conda-forge -c edipack edipack2
\end{lstlisting}
\noindent
which installs a bundle of Scifor and \NAME libraries together with
specific {\tt pkg-config} configurations files which can be used to
retrieve compilation and linking flags.

\subsection{OS Loading}\label{sSecInstallOSloading}
In oder to avoid possible conflicts or require administrative
privileges, the building step results get installed by default in a user home
directory, specified by the CMake variable {\tt PREFIX}.
In doing so, however, one misses the chance of automatic loading into
the operative system.

We offer different strategies to perform this action:
\begin{enumerate}
\item  A CMake generated configuration file for environment module
  which allows to load and unload the library at any time. This is
  preferred solution for HPC systems. 
\item A CMake generated bash script to be sourced (once or
  permanently) in any shell session to add \NAME library to the
  default environment.
\item A CMake generated pkg-config configuration file to be added in
  the pkg-config path itself.  
\end{enumerate}
An automatically generated recap message with all instructions is
generated at the end of the installation procedure.

>>>>

\section{Installation}\label{SecInstall}
The configuration and installation of \NAME is handled by CMake, which ensures
multi-platforms compabitility and dependencies resolution.  
The software builds into two distinct libraries.
The main one is a static Fortran library calledc{\tt  libedipack.a} which, alongside the compiled
modules, wraps the \NAME software in both its single- and multi-impurity variants.
A second dynamic library, {\tt libedipack\_cbinding.so}, together with the associated header
file{\tt edipack2\_cbinding.h}, enables interoperability with other programming languages.

\subsection{Structure}\label{sSecInstallStructure}
\NAME is a modular library organized into three main components. At its 
core, constituting the primary computational engine, 
is the exact diagonalization solver for the quantum single-impurity Hamiltonian.
Building on this, the library includes 
{\tt EDIpack2ineq}, an extension for handling multiple inequivalent 
impurity problems. Finally, the Fortran-C/C++ interface is provided for 
seamless integration with external software and the development of 
additional APIs.

@@

\subsection{Dependencies}\label{sSecInstallDependencies}
\NAME directly depends on two external libraries.
\begin{itemize}
\item {\bf SciFortran}: an open-source Fortran library to support
  mathematical and scientific software development. 
\item {\bf MPI} (optional): a distributed memory parallel communication layer with support to modern Fortran compiler.
\end{itemize}

@@

\subsection{Build and Install}\label{sSecInstallBuildInstall}
\subsubsection{Compilation from source}
The software can be installed from source as follows. The source code can
be retrieved directly from its GitHub repository, for instance using:
\begin{lstlisting}[style=mybash]
git clone https://github.com/edipack/EDIpack2.0 EDIpack2
\end{lstlisting}
Then, assuming to be in the software directory, a conventional
out-of-source building can be performed using two alternative compilation backends.

@@

\noindent
The CMake configuration can be further tuned using the following variables:
\begin{center}
\begin{tabular}{ l|l|l } 
 \hline
  {\bf Option}               & {\bf Scope} & {\bf Value} \\
  \hline
  -D{\bf PREFIX}          & Install directory  & $\sim$/opt/EDIpack2/TAG/PLAT/BRANCH\\
  -D{\bf USE\_MPI}       & MPI support  &  True/{\color{red}False}\\
  -D{\bf WITH\_INEQ}   & Inequivalent impurities support & {\color{red}True}/{False}\\
  -D{\bf VERBOSE}      & Verbose CMake output & {\color{red}True}/{False}\\ 
  -D{\bf BUILD\_TYPE} & Compilation flags & {\color{red}RELEASE}/TESTING/DEBUG/AGGRESSIVE \\
 \hline
\end{tabular}
\end{center}

The default target builds and install both the main library and the C-bindings.
Separate build targets for either are available. A recap message is printed at the end of the
CMake configuration step.

\subsubsection{Anaconda}
As an alternative we provide for both Linux and Mac Os systems
installation through Anaconda packages into a virtual
environment containing Python ($>3.10$).

The Conda package installation procedure reads:
\begin{lstlisting}[style=mybash]
conda create -n edipack
conda activate edipack
conda install -c conda-forge -c edipack edipack2
\end{lstlisting}
\noindent
which installs a bundle of SciFortran and \NAME libraries together with
specific {\tt pkg-config} configurations files, which can be used to
retrieve compilation and linking flags. In order to compile Fortran/C++ 
drivers, the {\tt compilers} conda package will need to be installed.

\subsection{OS Loading}\label{sSecInstallOSloading}
In order to avoid possible conflicts or the requirement for administrative
privileges, the building step results are installed by default in the user's {\tt HOME}
directory, specified by the CMake variable {\tt PREFIX}.
As a consequence, the system library and include paths will need to be updated by the user.

We offer different ways to perform this action:
\begin{enumerate}
\item  A CMake-generated configuration file for an environment module
  which allows to load and unload the library at any time. This is
  preferred solution for HPC systems. 
\item A CMake-generated bash script to be sourced (once or
  permanently) in any shell session to add \NAME library to the
  default environment.
\item A CMake generated {\tt pkg-config} file to be added in
  the {\tt PKG\_CONFIG\_PATH} itself.  
\end{enumerate}
An automatically generated recap message with all instructions is
generated at the end of the installation procedure.
